---
title: "Prep for LME  Monthly Averages"
author: "Just Berkhout"
date: "13-06-2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE)

library(rgdal)
library(RColorBrewer)

```

## Brief

> to spatially average the monthly fishmip inputs (pp intercept slopes, temperature et al) into 
Large marine ecosystems ?
I think there are shapefiles here: (also Reg has some if not) http://www.lmehub.net

## Goals
Averaging the fishmip inputs is the first step of three that will lead to integrating global fisheries catch input into the dbpm model. The steps are: 

1. **Provide fishmip inputs at the LME scale for model comparison and dbpm existing grid cell outputs**
1. To run and compare output of dbpm at the LME scale
1. To add catches into the model initially at LME scale (later grid cell)

The initial focus will be on LME. To complement the LME, FAO areas will then be used to cover the reamainder of global ocean area at similar scale.


## Large Marine Ecosystems
There is online interactive mapping of Large Marine Ecosystems (LME) on http://www.lmehub.net, but I couldn't find shape files or other spatial data in accessible form. I also couldn't find LME data on http://www.marineregions.org; the links are referencing http://www.lme.noaa.gov/, but resulting in a _404 Not Found_. The NOAA _does_ have an LME section at https://www.st.nmfs.noaa.gov/ecosystems/lme/index. It mentions that NOAA developed the Large Marine Ecosystem (LME) concept in cooperation with the University of Rhode Island over 30 years ago, and that there are 64 LMEs defined globally.

Because of the reference to the University of Rhode Island, I downloaded the LME shape file in zip archive from the Environmental Data Centre at the University of Rhode Island. Decsription here: http://lme.edc.uri.edu/index.php/digital-data/113-lme-polygon-boundaries and the archive itself here: http://lme.edc.uri.edu/images/Content/Downloads/DigitalData/LME66.zip. In this version of the LME data there are 66 (not 64) LME; presumably this reflects a more recent revision. 

Here I plot the bare LME polygon data in 12 colours. 

```{r plot LME shp, cache = TRUE}
LMEs <- readOGR(dsn ="LMEs66/", layer = "LMEs66", verbose = FALSE)
someColours = brewer.pal(11, "Spectral")
plot(LMEs, col=someColours[LMEs$LME_NUMBER%%12], border="black", lwd=1)

```

## Inventory of input files

The 'inputs files' used for the dbpm/Fishmip work are `.rds` files in `/rd/gem/private/fishmip_inputs/rds/`. They are organised in rcp scenarios, `rcp26`, `rcp45`, `rcp60`, `rcp85`. Each has 39570 files in them, corresponding with each ocean grid cell. The cells are ordinally numbered, repesent the surface of the earth in 1 degree grid cells, with all the land-based grid cells omitted.

the directory structure is given here:

```
fishmip_inputs/
  + misc
    + Seas and Submerged Lands Act 1973
      + Seas and Submerged Lands Act 1973.gdb
  + rds
    + rcp26
    + rcp45
    + rcp60
    + rcp85
```

``` {r init example input filename}
base_path <- "/rd/gem/private/"
file_name <- "fishmip_inputs/rds/rcp26/grid_9996_inputs2_ipsl-cm5a-lr_rcp26.rds"
```

An example path of one file is

> ``r file_name``

Each input .rds-file contains two data frames, called `depth` and `ts` respectively. 

The dataframe `depth` contains the metadata for the gridcell (`lon`, `lat`, `detph` and `gridnum`, where gridnum is the ordinal grid number that references the 360x180 earth grid as if the land grid cells had not been omitted). 

```{r inspect depth dataframe}

dat <- readRDS(sprintf("%s%s", base_path, file_name))

str(dat$depth)
```

The dataframe `ts` contains 21,645 timesteps for the variables `sst`, `sbt`, `er`, `intercept` and `slope`. The size and baseline of the timesteps are unclear atm, but they presumably contain a spinup, historical and projected component. The variables potentially correspond to the variables asked in the brief: `sst` Sea Surface Temperature, `sbt` Sea Bottom Temperature, `intercept` and `slope`. It is unclear how `er` maps to `pp`.



```{r inspect ts dataframe}
str(dat$ts)

```

### Inventory of fishmip outputs

If the `input files` in the brief refer to the __output files__ of the `dbpm` runs (and therefore to the __input files__ of the corresponding NetCDF files), then the inventory is different. The files are `.RData` files, are organised by rcp scenario, and several versions exist. The `aug_2017_2` run is the final one, that looks to have been translated into NetCDF, and therefore assumed the current production version. 

```
fishmip_outputs/
  + 20170802_trial
  + 20170802_trials2
  + aug_2017
    + netcdf
    + rcp26
    + rcp45
    + rcp60
    + rcp85
  + aug_2017_2
    + netcdf
    + netcdf2
      + decadal
    + netcdf2_trial
    + rcp26
    + rcp45
    + rcp60
    + rcp85
```

``` {r init example output filename}
base_path <- "/rd/gem/private/"
file_name <- "fishmip_outputs/aug_2017_2/rcp45/res_mts_agg_igrid_27601_ipsl-cm5a-lr_rcp45.RData"
```
Here i'm inspecting the file 

> ``r file_name``

This file contains a single data frame, called `agg`, that __aggregates__ 16 variables (13 variables + `lon`, `lat` and `depth`) into 5418 months (?) for each grid cell. It is not clear how any of these variables would relate to the variables "`pp`, `intercept`, `slopes`, `temperature` et al..." in the brief.

```{r inspect RData output file}

load(sprintf("%s%s", base_path, file_name))
str(agg)

```


## Method for Spatially Averaging

The grid cells appear to be represented by centre point (eg: the cell represented by lat=23.5 and lon=192.5, represents the grid cell bounded by minimum bounding rectangle mbr(24.0, 192.0 23.0, 193.0) ).

LME are exclusive (i.e. non overlapping contiguous polygonal areas)

grid cells belong to 0 or 1 LME

For each rcp scenario  
For each file (representing a grid cell) falling into an LME   
and only relevant timestep ranges selected (i.e no spinup)  
each 4th timestep value will be taken (to represent the monthly value)  
and aggregated into a data frame of schema:  

| Scenario | LME ID | Grid Num | ts/Month | sst | sbt | slope | intercept | pp/er |  
| ---------| ------ | -------- | -------- | --- | --- | ----- | --------- | ----- |

the resulting data frame will then be summarised by `mean` grouping by `Scenario`, `LME ID` and `ts/Month` resulting in a Monthly whole-of-area mean (average) of each of the variables for each of the LME.


```{r est df size, include=FALSE}

m <- matrix(1,nrow=(451*12*66*4),ncol=9)
m <- as.data.frame(m)
m[,1:2] <- sapply(m[,1:2],as.factor)

est_df_size <- gdata::humanReadable(object.size(m), standard = "SI", digits=0)


```

The output with three 'dimension' attributes (Scenario, LME ID and Timestep), and 5 variables, would consist of 1.4 Million rows (i.e 451 years of 12 months, across 66 LME and 4 Scenarios). This would be a dataframe of an estimated `r est_df_size`.



It should be possible to visualise these values in choropleth maps (fore each value) based on the polygons provided (and displayed above). It should be possible to visualise animates frame-by-frame monthly choropleth for displaying changes over time. 

## GROM - GRid cell Ordinal Mapping

* make mapping (ad find/develop language around ordinal grdcell numbering with or without the land omitted)

TO enable easier mapping between spatial data and the ocean only ordinal cell references by which the input files are referenced, I've made a mapping of gridnum to gridid. GridNum is the ordinal number of a cell in 1째x1째 grid starting from Latitude -89.5 and Longitude 0.5, running through to 89.5, 359.5, with all land and ocean included. Gridnum ranges from 1 to 64800. GridId is the contiguous ordinal numbering of all 1째x1째 grid cells that cover the ocean. GridID ranges from 1 to 39567.

Ive made a Grid cell Ordinal mapping (or GROM) that maps GridId (as found in the filenames of DBPM inputs) to GridNum as is found in the 1-row-dataframe `depth` in each inputfile. To enable spatial analysis and potentially depth-related analysis I have included the `lat` and `lon` and `depth` attributes in the mapping. The mapping is available as a file `grom.rds` in this repository, as is the code by which it was generated (`gridcell_ordinal_mapping.R`). 




## TO DO

* decide (and justify) LME66 vs LME64 (or accommodate both?)
* obtain LME and FAO areas (from Reg?)
* make mapping (ad find/develop language around ordinal grdcell numbering with or without the land omitted)
* are there multiple interpretations of "to spatially average the monthly fishmip inputs"
* dev averaging + for spatial averaging per run? per LME?
* visualisation for output
* in step 2 we'll run the model at LME scale; THe model can be run in aggregation mode (outputs similar to described above) or in detailed (non-aggregated) mode. Inspect what outputs that will generate, and what that is useful for. 

* one verification step along the way will be to view how averaged-grid-based-dbpm-outputs per LME compare with outputs obtained from LME scale (averaged) inputs. 

* the baseline method for including cells in an LME is if the cell __touches__ the LME, it's is include. 