---
title: "LME and FAO zones"
author: "Just Berkhout"
date: "10 July 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(raster)
library(dplyr)
```

## In short

Use the 0.5 degree grid with LME and FAO zones as used by Reg to aggregate the gridded input2 into LME scale inputs

Getting + inspecting the areas as provided by Reg in Areas.csv

```{r get areas, cache=TRUE}
areas <- read.csv("/rd/gem/private/fishmip_inputs/misc/LMEFAO_Areas/Areas.csv")

```

```{r inspect areas, echo=FALSE}

knitr::kable(head(areas))

```


Using attributes `LonDeg` and `LatDeg` to produce a grid with variable value `AreaCode`

```{r areas with ___Deg, cache=TRUE}

x <- raster(xmn=-180, xmx=180, ymn=-90, ymx=90, res=0.5, crs="+proj=longlat +datum=WGS84")
AreaGrid <- rasterize(areas[, c('LonDeg', 'LatDeg')], x, areas[, 'AreaCode'])
plot(AreaGrid, axes=FALSE, legend=FALSE)

```


Using `raster`, `rasterize` and the attibutes `LonCentre` and `LatCentre` to produce a grid with
variable `AreaCode` at input data resolution of 0.5°

``` {r area with ___Centre, cache=TRUE}
x2 <- raster(xmn=-180, xmx=180, ymn=-90, ymx=90, res=0.5, crs="+proj=longlat +datum=WGS84")
AreaGrid2 <- rasterize(areas[, c('LonCentre', 'LatCentre')], x2, areas[, 'AreaCode'])
plot(AreaGrid2, axes=FALSE, legend=TRUE)

```

Using `raster`, `rasterize` and the attibutes `LonCentre` and `LatCentre` to produce a grid with
variable `AreaCode` at resolution of 1°

``` {r area with ___Centre at resolution 1, cache=TRUE}
x3 <- raster(xmn=-180, xmx=180, ymn=-90, ymx=90, res=1, crs="+proj=longlat +datum=WGS84")
AreaGrid3 <- rasterize(areas[, c('LonCentre', 'LatCentre')], x3, areas[, 'AreaCode'])
dataType(AreaGrid3) <- "INT2S"
plot(AreaGrid3, axes=FALSE, legend=FALSE)

```

The datatype for this raster is `r dataType(AreaGrid3)`

```{r compare 1 degree rasters, cache=TRUE}

areas2 <- areas %>%
  filter(LonCentre - floor(LonCentre) == 0.75,
         LatCentre - floor(LatCentre) == 0.75) %>%
  mutate(LonCentre = LonCentre - 0.25,
         LatCentre = LatCentre - 0.25)

x4 <- raster(xmn=-180, xmx=180, ymn=-90, ymx=90, res=1, crs="+proj=longlat +datum=WGS84")
AreaGrid4 <- rasterize(areas2[, c('LonCentre', 'LatCentre')], x4, areas2[, 'AreaCode'])
dataType(AreaGrid4) <- "INT2S"
plot(AreaGrid4, axes=FALSE, legend=FALSE)



```

The datatype for this raster is `r dataType(AreaGrid4)`

```{r compare the two 1 degree rasters}
differentCells <- AreaGrid4%%AreaGrid3 != 0
plot(differentCells, axes=FALSE, legend=FALSE)
```

